---
title: 'Topic 7: Word Embeddings'
author: "Desik Somasundaram"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```


```{r packages, include = FALSE}
library(here)
library(tidytext)
library(tidyverse)
library(widyr)
library(irlba) #singluar value decomposition
library(broom) # creating search_synonym function
library(textdata)
library(ggplot2)
library(dplyr)
library(patchwork)

#https://semantle.com/
```

Today we are using climbing incident data from this repo: <https://github.com/ecaroom/climbing-accidents>. Some analysis (in Excel) on the data was written up into a Rock and Ice magazine article.

But I've constructed our data set (link below) by pulling a few key variables including the full text of each incident report.

```{r data,}
incidents_df<-read_csv("https://raw.githubusercontent.com/MaRo406/EDS_231-text-sentiment/825b159b6da4c7040ce8295b9eae2fbbe9991ffd/dat/climbing_report_text.csv")
```


```{r unigrams}
unigram_probs <- incidents_df %>%
    unnest_tokens(word, Text) %>%
    anti_join(stop_words, by = 'word') %>%
    count(word, sort = TRUE) %>%
    mutate(p = n / sum(n)) 
unigram_probs 
```


```{r}
skipgrams <- incidents_df %>%
    unnest_tokens(ngram, Text, token = "ngrams", n = 5) %>%
    mutate(ngramID = row_number()) %>% 
    tidyr::unite(skipgramID, ID, ngramID) %>%
    unnest_tokens(word, ngram) %>%
    anti_join(stop_words, by = 'word')

skipgrams
```

```{r}
#calculate probabilities
skipgram_probs <- skipgrams %>%
    pairwise_count(word, skipgramID, diag = TRUE, sort = TRUE) %>%
    mutate(p = n / sum(n))
```


```{r norm-prob}
#normalize probabilities
normalized_prob <- skipgram_probs %>%
    filter(n > 20) %>%
    rename(word1 = item1, word2 = item2) %>%
    left_join(unigram_probs %>%
                  select(word1 = word, p1 = p),
              by = "word1") %>%
    left_join(unigram_probs %>%
                  select(word2 = word, p2 = p),
              by = "word2") %>%
    mutate(p_together = p / p1 / p2)

#Which words are most associated with "rope"?   
normalized_prob %>% 
    filter(word1 == "rope") %>%
    arrange(-p_together)
```


```{r pmi}
pmi_matrix <- normalized_prob %>%
    mutate(pmi = log10(p_together)) %>%
    cast_sparse(word1, word2, pmi)    
 
#remove missing data
pmi_matrix@x[is.na(pmi_matrix@x)] <- 0
#run SVD using irlba() which is good for sparse matrices
pmi_svd <- irlba(pmi_matrix, 100, maxit = 500) #Reducing to 100 dimensions
#next we output the word vectors:
word_vectors <- pmi_svd$u
rownames(word_vectors) <- rownames(pmi_matrix)
```

```{r syn-function}
search_synonyms <- function(word_vectors, selected_vector) {
dat <- word_vectors %*% selected_vector
    
similarities <- dat %>%
        tibble(token = rownames(dat), similarity = dat[,1])

similarities %>%
       arrange(-similarity) %>%
        select(c(2,3))
}
```

```{r find-synonyms}
fall <- search_synonyms(word_vectors,word_vectors["fall",])
slip <- search_synonyms(word_vectors,word_vectors["slip",])
```

```{r plot-synonyms}
climbdata_plot <-slip %>%
    mutate(selected = "slip") %>%
    bind_rows(fall %>%
                  mutate(selected = "fall")) %>%
    group_by(selected) %>%
    top_n(15, similarity) %>%
    ungroup %>%
    mutate(token = reorder(token, similarity)) %>%
    ggplot(aes(token, similarity, fill = selected)) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~selected, scales = "free") +
    coord_flip() +
    theme(strip.text=element_text(hjust=0, size=12)) +
    scale_y_continuous(expand = c(0,0)) +
    labs(x = NULL, title = "Climbing data: What word vectors are most similar to slip or fall?")
         
```

```{r word-math}
snow_danger <- word_vectors["snow",] + word_vectors["danger",] 
search_synonyms(word_vectors, snow_danger)

no_snow_danger <- word_vectors["danger",] - word_vectors["snow",] 
search_synonyms(word_vectors, no_snow_danger)
```

### Assignment

Download a set of pretrained vectors, GloVe, and explore them. 

Grab data here:
<!-- download.file('<https://nlp.stanford.edu/data/glove.6B.zip>',destfile = 'glove.6B.zip')  -->
<!-- unzip('glove.6B.zip')  -->
<!-- Use this file: 'glove.6B.300d.txt' -->

1.  Recreate the analyses in the last three chunks (find-synonyms, plot-synonyms, word-math) with the GloVe embeddings. How are they different from the embeddings created from the climbing accident data? Why do you think they are different?

```{r}
library(data.table)
data_glove<-fread(here::here("dat","glove.6B",'glove.6B.300d.txt'), header = FALSE)
data_glove <- data_glove %>% 
  remove_rownames() %>% 
  column_to_rownames(var = "V1")
```

```{r find-synonyms2}

data_glove <- as.matrix(data_glove)

fall_2 <- search_synonyms(data_glove,data_glove["fall",])
slip_2 <- search_synonyms(data_glove,data_glove["slip",])
```

```{r plot-synonyms2}
glovedata_plot <- slip_2 %>%
    mutate(selected = "slip") %>%
    bind_rows(fall_2 %>%
                  mutate(selected = "fall")) %>%
    group_by(selected) %>%
    top_n(15, similarity) %>%
    ungroup %>%
    mutate(token = reorder(token, similarity)) %>%
    ggplot(aes(token, similarity, fill = selected)) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~selected, scales = "free") +
    coord_flip() +
    theme(strip.text=element_text(hjust=0, size=12)) +
    scale_y_continuous(expand = c(0,0)) +
    labs(x = NULL, title = "GloVe Data: What word vectors are most similar to slip or fall?")

climbdata_plot / glovedata_plot
         
```

The climbing data slip and fall histograms show that the word vectors caught climbing related terms such as accident, rope and line.
The GloVe data shows economic-related terms(prices, stocks, rates) for fall and cricket-related terms/players (Dravid, wicket, edged) for slip because slip is common fielding position in cricket. This is due to the GloVe data being more expansive than just climbing. 


```{r word-math2}
snow_danger <- data_glove["snow",] + data_glove["danger",] 
search_synonyms(data_glove, snow_danger)

no_snow_danger <- data_glove["danger",] - data_glove["snow",] 
search_synonyms(data_glove, no_snow_danger)
```
The GloVe data shows more weather and climate related incidents rather than climbing incidents in the addition equation and more macroeconomic related terms in the subtraction equation.



2.  Run the classic word math equation, "king" - "man" = ?
```{r word-math king-man}
king_man <- data_glove["king",] - data_glove["man",] 
search_synonyms(data_glove, king_man)

```


3.  Think of three new word math equations. They can involve any words you'd like, whatever catches your interest.
```{r word-math3}
batman_superman <- data_glove["batman",] + data_glove["superman",] 
search_synonyms(data_glove, batman_superman)

cricket_baseball <- data_glove["cricket",] - data_glove["baseball",] 
search_synonyms(data_glove, cricket_baseball)

air_water <- data_glove["air",] - data_glove["water",] 
search_synonyms(data_glove, air_water)

```